sensi_an_rotations1. Start of sensitivity analysis for rotations.

### 2018-07-06 inst/documents/rotations_notes.Rmd
for paper3

Just rough notes.

@@

*3/4/18*

~ set up initial sensi analysis of 1000 runs
~ 200 runs take < 30 mins

@@

*4/4/2018*

ooo with new ranges getting a few runs where there is a big advantage to rotation. Seems to be assoc with high dom_cos > 0.8
Only c 1% of scenarios, 2 out of 200

@@

*27/4/2018*

First draft of paper3

@@

# 30/4/18

100 runs with no cost all had identical times for rotations and sequences

@@

*2/5/18*

exchanging emails with Ian

@@

Before sending to Ian.
~ resize no square figs, make sure one per page

@@

@@

*today*
done ~ reread Ians text about dominance
~ move rotations repo back to my github
~ update R
done ~ tweet to malaria journal
~ think more about rotations datavis
done ~ read the alphey oxitech modelling papers

To talk with Ian about next week :
~ min_rwr_interval vs min_gens_switch_back

Look at Hackett R code. Could it be an even simpler model that we could use ?
The R code is here and looks quite accessible :
https://datadryad.org//resource/doi:10.5061/dryad.4s405

@@

~ try decision tree to predict rot_minus_seq from inputs

@@

~ *decide on min_rwr_interval for sensi analysis
~ one generation per month, 12 generations per year, 500 gens is ~40 years

@@

# good refs for paper
resistance : malERA2017_updated-research-agenda-resistance-in-malaria.pdf

@@

>>>>

@@

*3/4/18*
~ set up initial sensi analysis of 1000 runs
~ 200 runs take < 30 mins

@@

*4/4/2018*
ooo with new ranges getting a few runs where there is a big advantage to rotation. Seems to be assoc with high dom_cos > 0.8
Only c 1% of scenarios, 2 out of 200

@@

*27/4/2018*
First draft of paper3

@@

# 30/4/18
100 runs with no cost all had identical times for rotations and sequences

@@

*2/5/18*
exchanging emails with Ian

@@

Before sending to Ian.
~ resize non square figs, make sure one per page

@@

*21/5/2018*
Gould et al. Science paper.
Wicked evolution: Can we address the sociobiological dilemma of pesticide resistance?

If we are to address this recalcitrant issue of pesticide resistance, we must treat it as a “wicked
problem,” in the sense that there are social, economic, and biological uncertainties and complexities
interacting in ways that decrease incentives for actions aimed at mitigation. Here, we summarize
the interacting factors and conclude with a call for government support of ambitious landscape-level experiments to assess which pesticide use strategies decrease resistance risks.

In sexually reproducing weeds and insects, the rate of resistance evolution is strongly influenced by the relative fitness (dominant to recessive) of heterozygotes, and this sometimes depends on the dose of pesticide encountered in the field (Fig. 3). Thus, it is difficult (and controversial) to determine whether resistance is expected to evolve more rapidly to higher or lower application concentrations of a pesticide [e.g., 16, 17]. 
Even more complexity arises in attempts to predict resistance evolution when combinations of pesticides are applied (18, 19). Although the idea that such combinations will slow resistance evolution is theoretically controversial and lacks empirical support, mixtures are often recommended at the field level (15).

Model-based analysis has shown that if disease vector resistance to pyrethroids
becomes widespread, cases of malaria averted with ITNs could decline by 40% (25). Coupled with the estimate that bednets averted more than 65 million clinical malaria cases in sub- Saharan Africa in 2015 (7), and assuming that this figure provides a lower bound for potential cases averted in subsequent years, this would imply around 26 million additional clinical cases of malaria per year as a result of widespread vector resistance. Assuming an approximate lower bound cost of illness of at least $10 per malaria episode (26), insecticide resistance could conservatively cost sub-Saharan Africa at least $260 million per year.

*24/5/18*
replotted PRCC for all scenarios

resistance cost comes out as a negative impact on benefit of rotations. This could be because higher cost runs tend to be equal because they both sequences and rotations don't reach resistance thresholds. Maybe the PRCC is confusing ?

*25/5/18*
While out for a run I thought that the constraint on allele frequencies not going below their starting values could have implications for difference between rotations & sequences. I did 10k cost simulations with the constraint removed to check. Either way I should include something about this in the discussion.

*2/7/18* back to rotations after while on wos

try to knit rotations paper get : error in do.call ... object 'run_rot' not found 
Build & reload fixed it.

Also the citations aren't being picked up now ? Because I messed up export of library.bib from Mendeley, exporting just a single selected reference.

*4/7/2018* 
looking at Hackett2016, downloaded R code which is <100 lines and the script works to produce the first figure.

I was thinking that this could be a simpler model that we could also use with the same inputs that we use currently. This could possibly allow us to look at popn size, and to have a quick assessment of when resistance will increase and decrease.

The growth of the pest population is described by a density-dependent selection model (Roughgarden 1971) which uses the two parameter density dependence function developed by Maynard Smith & Slatkin (1973).

But maybe the way it works isn't appropriate for mosquitoes :

1. Pests divided between refuge and toxic patch
2. Pests within the Bt patch are exposed to toxins and mortality dependent on genotype.
3. Pest mortality via intraspecific competition (incl. costs of resistance)
4. Surviving larvae mature and mate at random to generate the next generation.

In mosquitoes competition likely to occur at larval stage, maybe doesn't matter ?

@@

~ check on the runs with the freq limit removed :
df_res_all_rot_cost10000_no_freq_limit.rda

~ think about whether I can implement allele frequencies just being constrained to their starting values the first time that each insecticide is used.

~ talk to Ian about Hackett2016 paper and whether we can use similarly simple model

~ ask Ian to check through the rotations algebra when he writes up the SI.

rotations paper :
~ develop flow of story
~ think about what questions managers will want answered

A) when do both strategies work (include looking at insecticides and rotation level)
 - maybe do PRCC
B) if neither work, when are rotations better ?

small changes :
~ dominance of resistance
~ how to term success. 'At least one insecticide still below resistance thresholds'

@@

@@

@@

# good refs for rotations paper
resistance : malERA2017_updated-research-agenda-resistance-in-malaria.pdf

@@

Slater2016. Use of an individual-based simulation model to explore and evaluate potential insecticide
resistance management strategies

### 2018-07-06 inst/documents/sensi_an_rotations1.Rmd
for paper3

    # for Ians Q1 set coverage to 1 & cost to 0 
    # April 2018, having go at including variable cost, may want to do some cost=0 runs later

    ## constant inputs ##
    linputs$coverage <- 1
    linputs$migration <- 0
    linputs$cost <- 0 #variable cost below
    linputs$max_gen <- max_gen

    #beware min_rwr_interval, it's to stop repeated small rotations
    linputs$min_rwr_interval = 1
    linputs$min_gens_switch_back <- 5
    #set plot to F if using grid.arrange as in phaps-rotations-better.Rmd
    linputs$plot <-     FALSE

    ## variable inputs ##
    #linputs$cost <-           runif(1, min=0,   max=0.1) 
    #linputs$cost <-           0 
    linputs$expo_hi <-        runif(1, min=0.4, max=0.9)  #min=0.1, max=0.9
    linputs$male_expo_prop <- runif(1, min=0,   max=1)
    linputs$eff <-            runif(1, min=0.5, max=1) # min=0.3, max=1
    linputs$rr  <-            runif(1, min=0.1, max=0.9)
    linputs$dom_sel <-        runif(1, min=0,   max=1)
    linputs$dom_cos <-        runif(1, min=0,   max=1) 
    linputs$start_freqs <-    runif(1, min=0.005,   max=0.1) #note not log yet   
    linputs$n_insecticides <- sample( 2:5, 1 )  #beware set replace=T if more than 1
    # dispersal (initially with cost set to 0)
    #linputs$coverage <-       runif(1, min=0.1, max=0.9)
    #linputs$migration <-      runif(1, min=0.1, max=0.9)

>>>>

    ## constant inputs ##
    linputs$coverage <- 1
    linputs$migration <- 0
    #linputs$cost <- 0 #variable cost below
    linputs$max_gen <- max_gen

    #beware min_rwr_interval, it's to stop repeated small rotations
    linputs$min_rwr_interval <- 5
    linputs$min_gens_switch_back <- 1
    linputs$no_r_below_start <- FALSE #BEWARE this has been set to TRUE in all earlier runs
    #set plot to F if using grid.arrange as in phaps-rotations-better.Rmd
    linputs$plot <-     FALSE

    ## variable inputs ##
    linputs$cost <-           runif(1, min=0,   max=0.1) 
    #linputs$cost <-           0 
    linputs$expo_hi <-        runif(1, min=0.4, max=0.9)  #min=0.1, max=0.9
    linputs$male_expo_prop <- runif(1, min=0,   max=1)
    linputs$eff <-            runif(1, min=0.5, max=1) # min=0.3, max=1
    linputs$rr  <-            runif(1, min=0.1, max=0.9)
    linputs$dom_sel <-        runif(1, min=0,   max=1)
    linputs$dom_cos <-        runif(1, min=0,   max=1) 
    linputs$start_freqs <-    runif(1, min=0.005,   max=0.1) #note not log yet   
    linputs$n_insecticides <- sample( 2:5, 1 )  #beware set replace=T if more than 1
    # dispersal (initially with cost set to 0)
    #linputs$coverage <-       runif(1, min=0.1, max=0.9)
    #linputs$migration <-      runif(1, min=0.1, max=0.9)
